### 2014-06-05 LifetableIncrement/LifetableIncrement.tex
already modified from first submission to DR, awaiting approval for revisions to be sent

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

\usepackage{amsmath}
\usepackage{caption}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{setspace}
\doublespacing
\bibpunct{(}{)}{,}{a}{}{;} 
\usepackage{url}
% for the d in integrals
\newcommand{\dd}{\; \mathrm{d}}
% end preamble.

\begin{document}

\title{The force of mortality by life lived is the force of increment by life
left in stationary populations}
\author{Author Redacted}
\maketitle

\section*{Structured abstract}

\subsection*{Background}
The age distribution and remaining lifespan distribution are identical in
stationary populations. The lifetable survival function is proportional to the
age distribution in stationary populations.

\subsection*{Objective}
We provide an alernative interpretation of the lifetable when viewed by
remaining years of life.

\subsection*{Conclusions}
The functions describing the mortality of birth cohorts over age are identical
to the functions describing the growth of death cohorts as time to death decreases in stationary populations.
\vspace{2cm}
\section*{Relationship}

Age can be defined as chronological (time since birth) or thanatological (time
until death). In a stationary population, decrement over
chronological age, as described by the lifetable, is
identical to increment over thanatological age.

\section*{Proof}

Chronological age structure equals thanatological age structure in a stationary
population \citep{brouard1989mouvements,vaupel2009life}:

\begin{equation}
\label{eq:vaupel}
c(a) = g(a) \quad\quad \text{,}
\end{equation}
where $c(a)$ is the stable chronological age structure expressed as a
proportion, and $g(a)$ is proportion of the population with $a$ remaining years
of life. Using $a$ to index chronological age and $y$ to index
thanatological age, the amount of time left until death. Equation
\eqref{eq:vaupel} is the same as:

\begin{equation}
\label{eq:equivalent}
c(a) = c^\star (y) \quad \text{for}\quad a = y \quad\quad\text{,}
\end{equation}
where $c^\star (y)$ refers to the stable thanatological age
structure. $c(a)$ is proportional to the survival
function, $l(a)$, which means that $l(a)$ is proportional to some
thanatological function, $l^\star (y)$, a stationary thanatological cumulative
\textit{increment} function. To untangle how this is so, it helps to be
explicit about what $l(a)$ represents. A birth cohort is a
group of individuals that are born at the same point in time and experience attrition over age until extinction, as described by $l(a)$. $l(a)$ is also
the sum of all deaths to the birth cohort at ages greater than $a$, and can be
thought of as the de-accumulation of the stock of future deaths over age.

\begin{equation}
l(a) = \int _{0}^\infty f(a+y) \dd y \quad \text{,}
\end{equation}

\noindent where $f(a)$ is the lifetable density function, often denoted by
$d(a)$. $f(a)$ gives the probability that a member of the birth cohort born in year $t$
will die in the year $t+a$. The deaths that will occur together in year
$t+a$ comprise a \textit{death cohort}. In the next year $t+1$,
$l(0)\cdot f(a-1)$ births increment to the $t+a$ death cohort in the stationary
population. Death cohorts grow monotonically in this way, starting with a few
members that will enjoy the maximum attainable lifespan, $ y = \omega$, and
accumulating new members as time approaches $t_0+\omega$ and $y$ decreases
toward 0:

\begin{equation}
l^\star (y) = \int _\omega^0 f(y) \dd y \quad \text{.}
\end{equation}

\noindent Extinction of the death cohort is simultaneous upon reaching
thanatological age 0. The lifetable deaths distribution is under this
perspective a distribution of the births to a death cohort, when read
from the highest to the lowest ages. The rate at which new births accumulate to
death cohorts over thanatological age, $\mu^\star (y)$ is given by
\begin{align}
\mu ^\star (y) &= \frac{-l^{\star^\prime} (y)}{l^\star (y)} \quad \text{,}
               \intertext{and since $l(a) = l^\star{a}$:}
               &= \frac{-l^\prime (y)}{l(y)} \notag\\
               &= \mu (y)
\end{align}

The rate of birth accumulation into death cohorts over thanatological age is
equal to the rate of death attrition of birth cohorts over chronological age in
stationary populations. The stock of deaths in the future for
chronological age $a$, $l(a)$, is symmetrically a description of births in the
past when structured by thanatological age, $l^\star(y)$. The remaining
lifetable columns are subject to similar reinterpretations when viewed under
thanatological age.

\section*{History and related results}

The term ``thanatological age'' has not previously appeared in the
literature, and was coined by Ken Wachter sometime prior to 2001. Thanatos was
the Greek god of death, which is used as the reference point for age under this
perspective. Explicit decompositions of chronological
age groups into remaining lifespan classes is to our knowledge only found in
\citet{brouard1986structure}, who redistributed
population pyramids by remaining years of life, and
\citet{miller2001increasing}, who examined medicare expenditure as a function of time until death using the
same lifetable identities. Equivalence of chronological and
thanatological age structures in stationarity, also known as Carey's equality,
is proven by \citet{brouard1989mouvements} and again in
 \citet{vaupel2009life}, which added detail to the more summary result reported
 in \citet{goldstein2009life}, that mean (or total) remaining lifespan is equal to mean (or total) life lived in the stationary population.
These latter two papers and \citet{goldstein2012historical} describe more of the
lineage of result presented here.

The estimation of remaining lifespan has motivated demography since
the invention of the lifetable, but this notion is typically dealt with as a
mean (expectancy). Age-specific remaining lifespan distributions are not often
used to explicitly decompose demographic quantities, such as population counts. Using
the above identities, one may estimate the total population with $y$
remaining years of life, $P(y)$ directly
\begin{equation}
\label{eq:getpy}
P(y) = \int _0 ^\infty P(a)\mu(a+y)\frac{l(a+y)}{l(a)} \dd a \quad \text{,}
\end{equation}
\noindent and so approximate a population's thanatological age structure
according to some mortality assumptions. This result differs from the more common approach
of summing the population within age classes bounded by some values interpolated
along the remaining life expectancy function (e.g.
\citet{sanderson2005average,sanderson2007new,sanderson2010remeasuring}, who
follow the line of \citet{hersch1944demographie} and \citet{ryder1975notes}).

In the present case of a stationary population, one decomposes back to birth
cohorts in much the same way as is evident from \eqref{eq:getpy}. The
probability that a member of death cohort $y'$ was born $a'$ years ago is equal to the probability that a member of birth cohort $a$ will die $y$ years in the future when $a = y'$ and $y = a'$.

\begin{align}
f(y | a) &= f(a' | y') \quad \text{for } a = y' \;\text{, } y = a'\\
&= \mu (a+y)\frac{l(a+y)}{l(a)}\\
&= \mu^\star (a'+y')\frac{l^\star(a'+y')}{l^\star(y')} \label{eq:redundant}
\end{align}

Equation \eqref{eq:redundant} seems redundant, but is less obvious when put into
words: The probability of dying $y$ years in the future given survival to chronological age
$a$ is the probability of surviving to chronological age $a+y$ given survival to
age $a$ times the force of mortality at age $a+y$, $\mu(a+y)$. Viewed thanatologically, the
probability of having been born $a'$ years in the past given that one has $y'$
remaining years of life is equal to the probability that someone in the
$y'$ death cohort was born more than $a'$ years ago given that they have already
been born times the force of increment at thanatological age $a'+y'$, $\mu
^\star (a'+y')$.

Similarly, taking $f(y|a)$ as the conditional distribution of remaining
lifespans, one may calculate the variance of age-specific remaining life
spans (thanatological age), where $e(a)$ is remaining life expectancy at age
$a$:
\begin{align}
e(a) =& \frac{\int _{y=0} ^\infty l(a+y) \dd y}{l(a)}
\intertext{The variance of $y$ given survival to age $a$ is}
\sigma ^2 (y|a) =& \int _{y=0}^\infty \left( e(a) - y\right)^2 f(y|a) \dd y
\quad \text{,}
\end{align}
\noindent and this function will have some non-monotonic pattern over age
(in human populations) that remains to be explored. In the reliability
literature, $\sigma ^2(y|a)$ is called the variance residual life function
(VRLF), and its properties have been described for various common distributions
\citep[see for example][]{gupta2006variance}.

\section*{Applications}
Thanatological age structure can be applied to stable populations (subject to a
growth rate, $r$), though we leave the description of a thanatological
renewal model for future work. Thanatological age equates individuals that share
a common terminal state rather than a common origin state. Here this is
the absorbing state of death, but the method generalizes to any terminal state
or lifecourse transition that can be modeled using lifetable techniques. Potential area
applications that may gain insights using such remaining-time methods
include morbidity, disability, late-life savings and investment behavior, or perhaps time to birth, menopause, retirement, or graduation.

\bibliographystyle{plainnat}
  \bibliography{bibliography}   % Use the BibTeX file ``References.bib''.

\end{document}

### 2014-06-06 LifetableIncrement/LifetableIncrement.tex
added line to copy compiled pdf to MainText.pdf
stuff for appending the cover page during review. header/footer text now in manuscript to save them time.

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% header, footers for review
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+
+%\chead{submitted as a relationship report to be included in the Special
+%Collection ``Formal Relationships'' edited by Joshua R. Goldstein and James W.
+%Vaupel}
+%\lfoot{Submission \#2518 for review}
+%\rfoot{}
+\fancyhf{}
+\fancyhead[C]{\footnotesize{submitted as a relationship report to be included in the
+Special Collection ``Formal Relationships'' edited by Joshua R. Goldstein and James W.
+Vaupel}}
+\fancyfoot[R]{\footnotesize{\bfseries{page \thepage}}}
+\fancyfoot[L]{\footnotesize\bfseries{Submission \#2518 for review}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrulewidth}{0pt}
+%\setlength\headheight{117.89105pt}
+\setcounter{page}{2}
+\pagestyle{fancy}
+
+\makeatletter
+\let\ps@plain\ps@fancy 
+\makeatother
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{amsmath}
@@
 \maketitle
-
 \section*{Structured abstract}
@@
 \end{equation}
-\noindent and so approximate a population's thanatological age structure
-according to some mortality assumptions. This result differs from the more common approach
+\noindent as \citet{brouard1986structure} did, and so approximate a population's
+thanatological age structure according to some mortality assumptions. This result differs from the more common approach
 of summing the population within age classes bounded by some values interpolated
@@
 
-Similarly, taking $f(y|a)$ as the conditional distribution of remaining
+Similarly, taking $f(y|a)$ as the conditional density of remaining
 lifespans, one may calculate the variance of age-specific remaining life
@@
 \end{document}
+% file.copy(from = "/home/triffe/git/LifetableIncrement/LifetableIncrement/LifetableIncrement.pdf", to = "/home/triffe/git/LifetableIncrement/LifetableIncrement/2518.pdf",overwrite= TRUE)

### 2014-06-06 LifetableIncrement/R/CoverPage.R

@@ -1,0 +1,7 @@
+# 1) copy compiled pdf to 2518.pdf
+
+# 2) append first page of pdf sent by Jana
+system("cd /home/triffe/git/LifetableIncrement/LifetableIncrement && pdftk cover.pdf LifetableIncrement.pdf cat output 2518.pdf")
+# one time only:
+#system("pdftk A=/home/triffe/Desktop/2518.pdf cat A1 output /home/triffe/git/LifetableIncrement/LifetableIncrement/cover.pdf")
+

### 2015-02-09 LifetableIncrement/ReplyReviewers.tex
added rao carey, Fries. Also reply to reviewers. main submission modified

\documentclass{article}

\usepackage{amsmath}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\newcommand{\dd}{\; \mathrm{d}}
\usepackage{url}
\begin{document}
\title{Reply to Reviewers of submission 2518: The force of mortality by life
lived is the force of increment by life left in stationary populations}
\author{Author Redacted}

Thanks for the useful comments, which improved the paper. I
hope that concerns are now met with the current revision.

Reviewer A gave math and notation suggestions, which were correct and which I
heeded. I think the primary concern is mended by changing former equation (4)
(now equation (5)) to:
\begin{equation}
l^\star (y) = \int _\omega^y f(t) \dd t \notag \quad \text{.} 
\end{equation}

Reviewer B suggested I add detail to the applications section. Admittedly, these
applications have to do with the thanatological age perspective in general
rather than the neatness of symmetry with respect to chronological decrement and
thanatological renewal within the lifetable. I hope my suggested applications
are an acceptable ersatz. The specific case I suggest is for the measurement of
morbidity patterns in order to approach the question of whether or not there
has been compression of morbidity. This question has been asked many times, and
it appears that the usual approach is to simply compare $e(65)$ with average
disability rates at ages $65+$. That is a very rough gauge. Measuring morbidity
in thanatological age would help disentangle morbidity change from lifespan change,
which is what we really want in order to say something about compression.

Reviewer C likened the thanatological perspective change to viewing the contents
of a room from above. This is a good analogy in the case of stationary
populations, but note that it only works for the marginal profile of lifetable
functions. The perspective change is more complex than that, bearing in mind
that the functions are decomposable into chronological and thanatological age
(described for the case of the pyramid). For an image of such heterogeneity, see
the colored images in this abstract to PAA, 2014: \url{http://paa2014.princeton.edu/papers/141036}
(before I knew about \citep{brouard1986structure}). I don't think an image is
necessary in this paper, to appreciate this, though, but I will oblige if the
wish is echoed further. I added a line at the end of the proof section ``The
main contribution of this relationship is to point out the symmery of chronological decrement and thanatological renewal processes in
the case of stationary populations.'' This contribution isn't revolutionary or
of great use, but the lifetable, stationary and stable populations, and the
renewal model stand at the core of demography, which ought to be
solid. I think tiny observations about them are worthy of publishing, and moreso
in this series, which values aesthetics. The actual main contribution of this
article is just to make a formal demographer smile.

Note that the stationary age structure finding, on which this result builds, has
appeared in the literatute again, since this paper was submitted
\citep{rao2014generalization}. These and previous authors were apparently also
not aware of Brouard's contributions. I hope to subtly set the record straight.

\bibliographystyle{plainnat}
  \bibliography{bibliography}

\end{document}

### 2015-02-09 LifetableIncrement/LifetableIncrement.tex
added rao carey, Fries. Also reply to reviewers. main submission modified

@@
 \fancyfoot[R]{\footnotesize{\bfseries{page \thepage}}}
-\fancyfoot[L]{\footnotesize\bfseries{Submission \#2518 for review}}
+\fancyfoot[L]{\footnotesize\bfseries{Submission \#2518, revised}}
 \renewcommand{\headrulewidth}{0pt}
@@
 
-Age can be defined as chronological (time since birth) or thanatological (time
-until death). In a stationary population, decrement over
+Age can be defined as chronological, $a$ (time since birth), or thanatological,
+$y$ (time until death). In a stationary population, decrement over
 chronological age, as described by the lifetable, is
-identical to increment over thanatological age.
+identical to increment over thanatological age. Define the chronological force
+of mortality, $\mu(a)$, and the thanatological force of increment,
+$\mu^\star(y)$. We have:
+\begin{equation}
+\mu(a) = \mu^\star(y)\quad\quad\text{for~}a=y\quad \text{.}
+\end{equation}
 
@@
 Chronological age structure equals thanatological age structure in a stationary
-population \citep{brouard1989mouvements,vaupel2009life}:
+population \citep{brouard1989mouvements,vaupel2009life,rao2014generalization}:
 
@@
 where $c(a)$ is the stable chronological age structure expressed as a
-proportion, and $g(a)$ is proportion of the population with $a$ remaining years
-of life. Using $a$ to index chronological age and $y$ to index
-thanatological age, the amount of time left until death. Equation
-\eqref{eq:vaupel} is the same as:
+proportion, and $g(a)$ is the proportion of the population with $a$ remaining
+years of life. Using $a$ to index chronological age and $y$ to index
+thanatological age (time left until death), equation \eqref{eq:vaupel} is the
+same as:
 
@@
 structure. $c(a)$ is proportional to the survival
-function, $l(a)$, which means that $l(a)$ is proportional to some
+function, $l(a)$, which means that $c^\star(y)$ is proportional to some
 thanatological function, $l^\star (y)$, a stationary thanatological cumulative
@@
 \begin{equation}
-l(a) = \int _{0}^\infty f(a+y) \dd y \quad \text{,}
+l(a) = \int _{0}^\omega f(a+t) \dd t \quad \text{,}
 \end{equation}
@@
 members that will enjoy the maximum attainable lifespan, $ y = \omega$, and
-accumulating new members as time approaches $t_0+\omega$ and $y$ decreases
+accumulating new members as time approaches $t+\omega$ and $y$ decreases
 toward 0:
@@
 \begin{equation}
-l^\star (y) = \int _\omega^0 f(y) \dd y \quad \text{.}
+l^\star (y) = \int _\omega^y f(t) \dd t \quad \text{.}
 \end{equation}
@@
 \mu ^\star (y) &= \frac{-l^{\star^\prime} (y)}{l^\star (y)} \quad \text{,}
-               \intertext{and since $l(a) = l^\star{a}$:}
+               \intertext{and since $l(a) = l^\star(a)$:}
                &= \frac{-l^\prime (y)}{l(y)} \notag\\
@@
 lifetable columns are subject to similar reinterpretations when viewed under
-thanatological age.
+thanatological age. The main contribution of this relationship is to point out
+the symmery of chronological decrement and thanatological renewal processes in
+the case of stationary populations.
 
@@
 thanatological age structures in stationarity, also known as Carey's equality,
-is proven by \citet{brouard1989mouvements} and again in
- \citet{vaupel2009life}, which added detail to the more summary result reported
- in \citet{goldstein2009life}, that mean (or total) remaining lifespan is equal to mean (or total) life lived in the stationary population.
+is proven by \citet{brouard1989mouvements}, again in
+ \citet{vaupel2009life},\footnote{\citet{vaupel2009life} added detail to the
+ more summary result reported in \citet{goldstein2009life}, which showed that
+ the mean (or total) remaining lifespan is equal to mean (or total) life lived in the stationary population.} and in more detail by \citet{rao2014generalization}.
 These latter two papers and \citet{goldstein2012historical} describe more of the
-lineage of result presented here.
+lineage of the result presented here.
 
@@
 
+The thanatological age perspective only offers the kind of profile symmetry
+presented in this paper for the theoretical case of stationary populations. For
+changing populations, the chronological and thanatological age perspectives
+typically offer different profiles of the same phenomena, due to changes in
+lifespan distributions and fluctuations in the birth flow, and therefore offer
+complementary information on population structure.
+
 \section*{Applications}
@@
 renewal model for future work. Thanatological age equates individuals that share
-a common terminal state rather than a common origin state. Here this is
-the absorbing state of death, but the method generalizes to any terminal state
+a common terminal state rather than a common origin state. In the present
+relationship, this is the absorbing state of death, but the method generalizes to any terminal state
 or lifecourse transition that can be modeled using lifetable techniques. Potential area
 applications that may gain insights using such remaining-time methods
-include morbidity, disability, late-life savings and investment behavior, or perhaps time to birth, menopause, retirement, or graduation.
+include morbidity, disability, late-life savings and investment behavior, or
+perhaps time to birth, menopause, retirement, or graduation. Glacial or open
+ice pack, old growth forest, and prison populations are other examples of
+aggregates for which remaining time structure is inherently of equal or greater
+interest than time passed. Populations of fixed or controlled size, or where
+entries are largely a function of exits are also prime candidates for analysis under some
+analogy to thanatological age. Examples of such populations include professional
+athletes in leagues, tenured professors, company directors, vehicle fleets.
+
+As a specific example, the question of morbidity compression has often been
+posed as a matter of comparing age of onset with remaining life expectancy
+\citep[e.g., ][]{fries2002aging,fries2003measuring}.\footnote{In the health
+literature, one often sees $e(65)$ compared with average disability levels above age 65.} In
+retrospect (within death cohorts), one can analytically separate between changes
+in lifespan and changes in morbidity as a function of time until death. Cross-tabulations of morbidity by chronological and thanatological age for two birth cohorts would allow the
+researcher to estimate changes in the morbidity profile over time free from
+distortion due to changes in the lifespan distribution, and so directly settle
+the question of morbidity compression.
 

### 2015-02-09 LifetableIncrement/ReplyReviewers.tex
reply to reviewers

@@
 \usepackage{url}
+
 \begin{document}
-\title{Reply to Reviewers of submission 2518: The force of mortality by life
-lived is the force of increment by life left in stationary populations}
+\title{Reply to Reviewers of submission 2518: ``The force of mortality by life
+lived is the force of increment by life left in stationary populations''}
 \author{Author Redacted}
-
+\maketitle
 Thanks for the useful comments, which improved the paper. I
@@
 thanatological renewal within the lifetable. I hope my suggested applications
-are an acceptable ersatz. The specific case I suggest is for the measurement of
+are an acceptable ersatz. The specific case I now add is for the measurement of
 morbidity patterns in order to approach the question of whether or not there
@@
 solid. I think tiny observations about them are worthy of publishing, and moreso
-in this series, which values aesthetics. The actual main contribution of this
-article is just to make a formal demographer smile.
+in this series, which values aesthetics.
 

### 2015-02-09 LifetableIncrement/R/CoverPage.R
modified file path for new laptop

@@ -1,7 +1,9 @@
 # 1) copy compiled pdf to 2518.pdf
 
 # 2) append first page of pdf sent by Jana
-system("cd /home/triffe/git/LifetableIncrement/LifetableIncrement && pdftk cover.pdf LifetableIncrement.pdf cat output 2518.pdf")
+system("cd /home/tim/git/LifetableIncrement/LifetableIncrement && pdftk cover.pdf LifetableIncrement.pdf cat output 2518.pdf")
+# old laptop:
+#system("cd /home/triffe/git/LifetableIncrement/LifetableIncrement && pdftk cover.pdf LifetableIncrement.pdf cat output 2518.pdf")
 # one time only:
-#system("pdftk A=/home/triffe/Desktop/2518.pdf cat A1 output /home/triffe/git/LifetableIncrement/LifetableIncrement/cover.pdf")
+#system("pdftk A=/home/tim/Desktop/2518.pdf cat A1 output /home/tim/git/LifetableIncrement/LifetableIncrement/cover.pdf")
 

### 2015-03-05 LifetableIncrement/LifetableIncrement.tex
in final part of demonstration, changed a's to y's for clarity
added Chiang reference. Modified based on reviewer A suggestions.

@@
 \begin{equation}
-l(a) = \int _{0}^\omega f(a+t) \dd t \quad \text{,}
+\label{eq:la}
+l(a) = \int _{0}^{\omega-a} f(a+t) \dd t \quad \text{,}
 \end{equation}
@@
 $d(a)$. $f(a)$ gives the probability that a member of the birth cohort born in year $t$
-will die in the year $t+a$. The deaths that will occur together in year
-$t+a$ comprise a \textit{death cohort}. In the next year $t+1$,
-$l(0)\cdot f(a-1)$ births increment to the $t+a$ death cohort in the stationary
-population. Death cohorts grow monotonically in this way, starting with a few
-members that will enjoy the maximum attainable lifespan, $ y = \omega$, and
-accumulating new members as time approaches $t+\omega$ and $y$ decreases
-toward 0:
+will die in the year $t+a$. Since $t+a$ refers to the future,
+let us switch to index $y$, for thanatological age. Under
+stationarity, $f(y)$ is also the probability at birth that an individual born
+$y$ years ago will die in present year $t$. The deaths that occur together in
+year $t$ comprise a \textit{death cohort}.
+In year, $t-1$, $l(0)\cdot f(1)$ births increment to the year $t$ death cohort in the stationary population.
+Death cohorts grow monotonically in this way, starting
+with a few members that will enjoy the maximum attainable lifespan, $ y =
+\omega$, i.e., born $\omega$ years ago in year $t-\omega$ and then expiring in
+year $t$. New births accumulate into a given death cohort as $y$ decreases from
+$\omega$ toward thanatological age 0. From the vantage point of year
+$t$, we define the year $t+y$ death cohort ($y$ years in the future), $l^\star
+(y)$ as the members born up to year $t$ that will die in exactly $y$ years.
 
 \begin{equation}
-l^\star (y) = \int _\omega^y f(t) \dd t \quad \text{.}
+\label{eq:lstar}
+l^\star (y) = \int _{-(\omega - y)}^0 f(y-t) \dd t \quad \text{.}
 \end{equation}
@@
 \noindent Extinction of the death cohort is simultaneous upon reaching
-thanatological age 0. The lifetable deaths distribution is under this
-perspective a distribution of the births to a death cohort, when read
-from the highest to the lowest ages. The rate at which new births accumulate to
-death cohorts over thanatological age, $\mu^\star (y)$ is given by
+thanatological age 0, $y$ years in the future. Equation~\ref{eq:lstar} is equal
+to \ref{eq:la}:
+
+\begin{equation}
+\begin{split}
+l^\star (y) = \int_{-(\omega-y)}^0 f(y-t)\dd t = \int _{-\omega}^{-y}
+f(-t)\dd t = \int _y ^\omega f(t) \dd t = l(y) \quad \text{.}
+\end{split}
+\end{equation}
+
+The lifetable deaths distribution is under
+this perspective a distribution of the births to a death cohort, when read
+from the highest to the lowest chronological ages. The rate at which new births
+accumulate to death cohorts over thanatological age, $\mu^\star (y)$, is given
+by
 \begin{align}
 \mu ^\star (y) &= \frac{-l^{\star^\prime} (y)}{l^\star (y)} \quad \text{,}
-               \intertext{and since $l(a) = l^\star(a)$:}
-               &= \frac{-l^\prime (y)}{l(y)} \notag\\
-               &= \mu (y)
+               \intertext{and since $l(a) = l^\star(y)$ for $a = y$,}
+               &= \frac{-l^\prime (a)}{l(a)} \notag\\
+               &= \mu (a) \quad \text{.}
 \end{align}
 
-The rate of birth accumulation into death cohorts over thanatological age is
-equal to the rate of death attrition of birth cohorts over chronological age in
-stationary populations. The stock of deaths in the future for
-chronological age $a$, $l(a)$, is symmetrically a description of births in the
+The rate of birth accumulation into death cohorts over thanatological age,
+$\mu^\star (y)$, is equal to the rate of attrition of birth cohorts
+over chronological age, $\mu (a)$ in stationary populations. The stock of deaths in
+the future for chronological age $a$, $l(a)$, is symmetrically a description of births in the
 past when structured by thanatological age, $l^\star(y)$. The remaining
@@
 thanatological age. The main contribution of this relationship is to point out
-the symmery of chronological decrement and thanatological renewal processes in
+the symmetry of chronological decrement and thanatological renewal processes in
 the case of stationary populations.
@@
 \noindent and this function will have some non-monotonic pattern over age
-(in human populations) that remains to be explored. In the reliability
-literature, $\sigma ^2(y|a)$ is called the variance residual life function
-(VRLF), and its properties have been described for various common distributions
+(in human populations) that remains to be explored.\footnote{This
+definiton works out to be the same as \citet{chiang1984life}, Chapter 10,
+Equation 6.10.} In the reliability literature, $\sigma ^2(y|a)$ is called the variance residual life function (VRLF), and its properties have been described for various common distributions
 \citep[see for example][]{gupta2006variance}.
@@
 analogy to thanatological age. Examples of such populations include professional
-athletes in leagues, tenured professors, company directors, vehicle fleets.
+athletes in leagues, tenured professors, company directors, and vehicle fleets.
 

### 2015-03-05 LifetableIncrement/ReplyReviewer2.tex
commit reply 2 to reviewers, and minro fixes to manuscript

\documentclass{article}

\usepackage{amsmath}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\newcommand{\dd}{\; \mathrm{d}}
\usepackage{url}

\begin{document}
\title{Second reply to reviewers of submission 2518: ``The force of mortality by
life lived is the force of increment by life left in stationary populations''}
\author{Author Redacted}
\maketitle

My thanks to the reviewers for the corrections and constructive suggestions.
Reviewer A found an error that owes to my own misunderstanding of integrals,
which until now I've naively understood as perfectly analogous to discrete sums.
In other words, the old equation (5) works when you draw it on paper, but now I
understand why it does not work with calculus notation. Reviewer A also
provided a solution that saves the argument, and for which I am grateful.

I've only made yet another innocuous change to both my own equation (4) and
reviewer A's suggestion for equation (5):

Equation 4 Was:
\begin{equation}
\label{eq:la}
l(a) = \int _{0}^{\omega} f(a+t) \dd t \quad \text{,}
\end{equation}
Now:
\begin{equation}
\label{eq:la}
l(a) = \int _{0}^{\omega-a} f(a+t) \dd t \quad \text{.}
\end{equation}
This is just aesthetic, I suppose, but really it's the sum ($a+t$)
that takes a maximum value of $\omega$. It's an innocuous change since higher
values are all zero. I made this change because it helped me work through reviewer A's
equation of $l^\star (y)$ with $l(a)$ when $a=y$, so I changed the reviewers
alternative definition of $l^\star (y)$ to:
\begin{equation}
\label{eq:lstar}
l^\star (y) = \int _{-(\omega - y)}^0 f(y-t) \dd t \quad \text{,}
\end{equation}
and everything works out to be the same by my reckoning.

The text between equations 4 and 5 has also been substantially changed to
better reflect the new formulation of $l^\star (y)$, and I hope it is clear.

I also changed the indexing of the final result, and included ``for $a=y$'', as
requested by the reviewer. You're right, it's easier to understand that way.
-----------------------------

Reviewer C looked for the old Brouard material. This is a link a a 1989 version of the Brouard paper with a version of Carey's inequality in it:

\url{http://sauvy.ined.fr/brouard/enseignements/iford/mouvementetmodeles.pdf}

The juicy part is section 4.2.2, ``Pyramide des ann\'{e}es a vivre'', which is
couched in terms of the stationary population. I do think that he will have
first written this down earlier, since he cites the relationship in
\citet{brouard1986structure}, which is better known for having introduced the
measure 'CAL'.

Here's a link to the 1986 paper: 
\url{http://www.persee.fr/web/revues/home/prescript/article/espos_0755-7809_1986_num_4_2_1120}

The key sentence in the  1986 paper, which led me to discover the other more
obscure source, is at the bottom of page 159, and continues into page 161 (pg
160 is a figure page)- he states Careys equality in plain French and cites a 1985 manuscript ``Mouvements et mod\`{e}les de population''
to back it up. I've never been able to locate the 1985 version, but it appears
to be something he maintained over a series of years, and the
first linked manuscript above appears to be from 1989 (so I cite it as such). Looking at the inside title page it says 1986. It appears to have been placed online in 2002.

To settle the date more exactly, one would have to ask N. Brouard when he
thought of it (or where he saw it). I imagine he went to Cameroon in 1985 with a
notebook of concepts to demonstrate, and either he knew it beforehand, or it occurred to him there, but it wasn't, as far as I
can tell, in a distributed print form until 1989. He apparently cited his 1985
workshop materials, though I could be mistaken. I've not seen a hard copy.

Regarding \citet{rao2014generalization}, the citation seems OK to me, but it was
easy to miss since the bibliography is alphabetical by last name but printed
starting with first name (Arni). I'll reformat as the journal requires.
-----------------------------------
Finally, I added a reference to \citet{chiang1984life}, who provides an
alternative but equivalent expression of $\sigma^2(y|a)$ (in the History \&
Related Results section), and I also fixed some minor typos here and there.

Thanks for the careful reviewing.

\bibliographystyle{plainnat}
  \bibliography{bibliography}

\end{document}

### 2015-03-05 LifetableIncrement/LifetableIncrement.tex
commit reply 2 to reviewers, and minro fixes to manuscript

@@
 \subsection*{Objective}
-We provide an alernative interpretation of the lifetable when viewed by
+We provide an alternative interpretation of the lifetable when viewed by
 remaining years of life.
@@
 year $t$ comprise a \textit{death cohort}.
-In year, $t-1$, $l(0)\cdot f(1)$ births increment to the year $t$ death cohort in the stationary population.
+In year $t-1$, $l(0)\cdot f(1)$ births increment to the year $t$ death cohort in the stationary population.
 Death cohorts grow monotonically in this way, starting

### 2015-03-05 LifetableIncrement/ReplyReviewer2.tex
final commit for this round of revisions

@@
 My thanks to the reviewers for the corrections and constructive suggestions.
-Reviewer A found an error that owes to my own misunderstanding of integrals,
-which until now I've naively understood as perfectly analogous to discrete sums.
+\textbf{Reviewer A} found an error that owes to my own misunderstanding of
+integrals, which until now I've naively understood as perfectly analogous to discrete sums.
 In other words, the old equation (5) works when you draw it on paper, but now I
@@
 requested by the reviewer. You're right, it's easier to understand that way.
------------------------------
 
-Reviewer C looked for the old Brouard material. This is a link a a 1989 version of the Brouard paper with a version of Carey's inequality in it:
+\textbf{Reviewer C} looked for the old Brouard material. This is a link a a
+1989 version of the Brouard paper with a version of Carey's inequality in it:
 
@@
 starting with first name (Arni). I'll reformat as the journal requires.
------------------------------------
-Finally, I added a reference to \citet{chiang1984life}, who provides an
-alternative but equivalent expression of $\sigma^2(y|a)$ (in the History \&
+
+\textbf{Finally}, I added a reference to \citet{chiang1984life}, who provides
+an alternative but equivalent expression of $\sigma^2(y|a)$ (in the History \&
 Related Results section), and I also fixed some minor typos here and there.
 
-Thanks for the careful reviewing.
+Thanks to all for the careful reviewing
 

### 2015-04-02 LifetableIncrement/LifetableIncrement.tex
moved sentence, changed integration limits to make less redundant

@@
 %%There is just one section and two subsections.
-\documentclass{article}
+\documentclass[leqno]{article}
+% leqno is used for equations on the left.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%% To the editors: this part of the preamble was used for markup
+%%%%%%%%%% during the review process, but I comment it out now so you have
+%%%%%%%%%% an easier time of editing
 % header, footers for review
-\usepackage{fancyhdr}
-\pagestyle{fancy}
+%\usepackage{fancyhdr}
+%\pagestyle{fancy}
 
@@
 %\rfoot{}
-\fancyhf{}
-\fancyhead[C]{\footnotesize{submitted as a relationship report to be included in the
-Special Collection ``Formal Relationships'' edited by Joshua R. Goldstein and James W.
-Vaupel}}
-\fancyfoot[R]{\footnotesize{\bfseries{page \thepage}}}
-\fancyfoot[L]{\footnotesize\bfseries{Submission \#2518, revised}}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}
+%\fancyhf{}
+%\fancyhead[C]{\footnotesize{submitted as a relationship report to be included
+% in the Special Collection ``Formal Relationships'' edited by Joshua R. Goldstein and James W.
+%Vaupel}}
+%\fancyfoot[R]{\footnotesize{\bfseries{page \thepage}}}
+%\fancyfoot[L]{\footnotesize\bfseries{Submission \#2518, revised}}
+%\renewcommand{\headrulewidth}{0pt}
+%\renewcommand{\footrulewidth}{0pt}
 %\setlength\headheight{117.89105pt}
-\setcounter{page}{2}
-\pagestyle{fancy}
+%\setcounter{page}{2}
+%\pagestyle{fancy}
 
-\makeatletter
-\let\ps@plain\ps@fancy 
-\makeatother
+%\makeatletter
+%\let\ps@plain\ps@fancy 
+%\makeatother
 
@@
 \usepackage{graphicx}
-\usepackage{natbib}
+\usepackage{natbib} % natbib called for bibtex, used with defaults. 
+                    % if formatting is eventually done using package arguments,
+                    % it'd be helpful for authors to know the combo of arguments
+                    % required to meet reference guidelines. Or maybe a style
+                    % could be written?
 \usepackage{setspace}
+\usepackage{authblk} % for affiliation
 \doublespacing
@@
 left in stationary populations}
-\author{Author Redacted}
+\author{Tim Riffe}
+\affil{Department of Demography, University of California, Berkeley}
 \maketitle
@@
 lifetable columns are subject to similar reinterpretations when viewed under
-thanatological age. The main contribution of this relationship is to point out
-the symmetry of chronological decrement and thanatological renewal processes in
-the case of stationary populations.
+thanatological age.
 
@@
 
-The term ``thanatological age'' has not previously appeared in the
-literature, and was coined by Ken Wachter sometime prior to 2001. Thanatos was
+The main contribution of this relationship is to point out
+the symmetry of chronological decrement and thanatological renewal processes in
+the case of stationary populations. The term ``thanatological age'' has not
+previously appeared in the literature, and was coined by Ken Wachter sometime prior to 2001. Thanatos was
 the Greek god of death, which is used as the reference point for age under this
@@
 of summing the population within age classes bounded by some values interpolated
-along the remaining life expectancy function (e.g.
+along the remaining life expectancy function (e.g.,
 \citet{sanderson2005average,sanderson2007new,sanderson2010remeasuring}, who
@@
 \begin{align}
-e(a) =& \frac{\int _{y=0} ^\infty l(a+y) \dd y}{l(a)}
+e(a) =& \frac{\int _{0} ^\infty l(a+y) \dd y}{l(a)}
 \intertext{The variance of $y$ given survival to age $a$ is}
-\sigma ^2 (y|a) =& \int _{y=0}^\infty \left( e(a) - y\right)^2 f(y|a) \dd y
+\sigma ^2 (y|a) =& \int _{0}^\infty \left( e(a) - y\right)^2 f(y|a) \dd y
 \quad \text{,}
@@
 
+\section*{Acknowledgements}
+A big thank you to three anonymous reviewers for providing helpful 
+suggestions, comments, and corrections. This work was supported by the National Institute On
+Aging of the U.S. National Institutes of Health (NIH) under Award Numbers
+R01-AG011552 and R01-AG040245. The content is solely the responsibility of the
+author and does not necessarily represent the official views of the NIH.
+
 \bibliographystyle{plainnat}

### 2015-04-23 LifetableIncrement/LifetableIncrement.tex
final changes to manuscript

@@
 Chronological age structure equals thanatological age structure in a stationary
-population \citep{brouard1989mouvements,vaupel2009life,rao2014generalization}:
+population \citep{brouard1989mouvements,vaupel2009life}:
 
@@
 the case of stationary populations. The term ``thanatological age'' has not
-previously appeared in the literature, and was coined by Ken Wachter sometime prior to 2001. Thanatos was
-the Greek god of death, which is used as the reference point for age under this
+previously appeared in the literature, and was coined by Kenneth Wachter
+sometime prior to 2001. Thanatos was the Greek god of death, which is used as the reference point for age under this
 perspective. Explicit decompositions of chronological
@@
 is proven by \citet{brouard1989mouvements}, again in
- \citet{vaupel2009life},\footnote{\citet{vaupel2009life} added detail to the
+ \citet{vaupel2009life}.\footnote{\citet{vaupel2009life} added detail to the
  more summary result reported in \citet{goldstein2009life}, which showed that
- the mean (or total) remaining lifespan is equal to mean (or total) life lived in the stationary population.} and in more detail by \citet{rao2014generalization}.
-These latter two papers and \citet{goldstein2012historical} describe more of the
+ the mean (or total) remaining lifespan is equal to mean (or total) life lived in the stationary population.} 
+This latter paper and \citet{goldstein2012historical} describe more of the
 lineage of the result presented here.
@@
 entries are largely a function of exits are also prime candidates for analysis under some
-analogy to thanatological age. Examples of such populations include professional
+variant of thanatological age. Examples of such populations include professional
 athletes in leagues, tenured professors, company directors, and vehicle fleets.
